Testing and Isolation Efficacy; Insights from a Simple Epidemic Model

### 2020-11-19 SIR_manuscript.tex
version 0

 [Fady?] 
 Something needs to be said here about simulation part which was done by using R, using the expression on $\Rnum$ and not the approximation, which give the panel plot in Figure \ref{pan}.
The analytical calculation of the next generation matrix, $G$, and the derivation of $\Rnum$ was carried out in Maple (ref) by using simple linear Algebra package. The numerical simulation of $\Rnum$ in the specified parameter space and plotting the contours of $\Rnum$ was carried out in R by using ggplot (ref?). It is notable that in the simulation part, the explicit expression of $\Rnum$ was used and not the approximation \eqref{eq:R0appr}.

@@

We also carried out simulation to support the analytical results in Proposition \ref{prop1}.

[comments about the figure:]
1- use a single legend, change the range manually.
\begin{figure}[h!]
\centering
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{./pix/R0contour_random.png}
\end{subfigure}
%
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{./pix/R0contour_TTI.png}
\end{subfigure}
\caption{Behaviour of the basic reproduction number, $\Rnum$, with respect to changes in the underlying parameters; remind the reader about the parameters and their values in this simulations.}
\label{pan}
\end{figure}

When random testing regime is applied, $W_S=W_I=W_R$,

(i) in the case of perfect isolation of the reported individuals, $\eta_t=0$, the followings are inferred (see the first few columns of the panel plot); 
 - variations of the $\eta_w$ has a negligible effect on the critical testing.
 - for low $\omega$, $R_0$ increases as $\omega$ increases but saturates. This means that when the reporting process is very very slow, increasing the speed of the reporting causes higher $R_0$. Maybe because more people will move to $I_u$ [?, not very sure here.]
 (ii) in case of the non-perfect isolation (see the last column), when $\eta_w$ increases (more awaiting people follow the isolation rules) faster reporting reduces $R_0$.  
Note that it is sensible to expect that $\partial{R_0}/\partial{\omega}<0$. Here we showed that this can be less true. 
%%%%%%%%%%%%
[Fady: this subsection and the next belong to the Appendix since they are very detailed. Would you 
(i) clean up the writing, there are some repeated expressions that can be referenced, (ii) write a paragraph for each to reflect what the message/application is?]
\subsection{rate of returning tests}
The linearization of $\Rnum$ around $\rho=0$ is
\begin{equation}\label{linearization}
\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big). 
\end{equation}

This inequality quantifies the exact relationships between model parameters that result in returning tests more rapidly being favorable. To give a biological interpretation, we describe the qualitative \textit{trends} predicted by the inequality. Returning test results more rapidly \textit{tends} to be more favorable when

@@

@@

>>>>

 [Fady?] 
 Something needs to be said here about simulation part which was done by using R, using the expression on $\Rnum$ and not the approximation, which give the panel plot in Figure \ref{pan}.
The analytical calculation of the next generation matrix, $G$, and the derivation of $\Rnum$ was carried out in Maple (ref) by using simple linear Algebra package. The numerical simulation of $\Rnum$ in the specified parameter space and plotting the contours of $\Rnum$ was carried out in R by using ggplot (ref?). It is notable that in the simulation part, the explicit expression of $\Rnum$ was used and not the approximation \eqref{eq:R0appr}. This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $W_S=W_I=W_R$, and (ii) non-random testing, when testing weight are not equal. For simmulation purposes we chose $W_S=W_I=W_R=1$ for random testing, and $W_S=0.3$ and $W_I=W_R=1$ for non-random testing strategy. Note that the critical contour of $\Rnum=1$ is plotted in solid line in Figure \ref{pan}.

@@

[comments about the figure:]
1- use a single legend, change the range manually.
\begin{figure}[h!]
\centering
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{./pix/R0contour_random.png}
\caption{}\label{p.a}
\end{subfigure}
%
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{./pix/R0contour_TTI.png}
\caption{}\label{p.b}
\end{subfigure}
\caption{Behaviour of the basic reproduction number, $\Rnum$, with respect to changes in the underlying parameters; remind the reader about the parameters and their values in this simulations. the solid line in each panel represents the critical contour of $\Rnum=1$.}
\label{pan}
\end{figure}

Our numerical simulation in support of the analytical results in Proposition \ref{prop1} are presented in Figure \ref{pan}. It can be infered that when random testing strategy is applied, Figure \ref{p.a}, comparing to the non-random testing strategy, Figure \ref{p.a}, the critical testing intensity of the coresponding panels are lower in non-random testing. Also, it appears that speeding the test reporting, i.e., increasing $\omega$, does not significantly lower $\Rnum$ when awaiting people follow the isolation perfectly, i.e., $\eta_w$ is closer to 0. However, speeding the test reporting reduces the epidemy more when the awaiting people less follow the isolation, i.e., $\eta_w$ is closer to 1.

Further, we derived an expression that quantifies the exact relationships between model parameters that result in returning tests more rapidly being favorable \eqref{eq:necsuf}. To give a biological interpretation, we describe the qualitative \textit{trends} predicted by the inequality. Returning test results more rapidly \textit{tends} to be more favorable when

@@

[Fady, would you sumerize the results of "Expensive vs. cheap tests" into a paragraph and leave the Mathy stuff in the Appendix.]
%%%%%%%%%%%%

@@

% %%%%%%%
[Fady: this subsection and most of the next belong to the Appendix since they are very detailed. Would you clean up the writing, there are some repeated expressions that can be referenced?]

\subsection{rate of returning tests}
The linearization of $\Rnum$ around $\rho=0$ is
\begin{equation}\label{linearization}
\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big). 
\end{equation}
